### 2022-10-13 doc/manuscript/translocation.qmd
Add additional details to journal summary
Add methods for population viability modeling

@@

Science:

@@

Nature

@@

PNAS

@@

Nature Communications

@@

    -   Papers published by the journal represent important advances of significance to specialists within each field.

@@

>>>>

**Model description**

To determine the implications of acquired Bd resistance on long-term population viability, we developed a dynamic, population model for MYLF. Our central question was: how does the magnitude and variation in observed adult survival probability across translocated populations affect the long-term persistence probability of populations? We developed a model that tracked seven state variables of an MYLF population: the density of translocated adults ($A_T$), the density of adults naturally recruited into the population ($A_R$), the density of first year larvae ($L_1$), the density of second year larvae ($L_2$), the density of third year larvae ($L_3$), the density of first year juveniles ($J_1$), and the density of second year juveniles ($J_2$). We divided adults into two classes $A_T$ and $A_R$ because there was evidence that the survival probability of translocated adults and naturally recruited adults differed.

We modeled the dynamics of these seven state variables using a discrete-time, stage-structured model where a time step is one year. The dynamics are given by

$$
\begin{bmatrix}
L_1 \\
L_2 \\
L_3 \\
J_1 \\
J_2 \\
A_R \\ 
A_T
\end{bmatrix}(t + 1) = 
\begin{bmatrix}
  0 & 0 & 0 & 0 & 0 & \sigma_{A_R} p_F F & \sigma_{A_T} p_F F \\
  \sigma_{L1} p_{L1} & 0 & 0 & 0 & 0 & 0 & 0 \\
  0 & \sigma_{L2} p_{L2} & 0 & 0 & 0 & 0 & 0\\
  \sigma_{L1} (1 - p_{L1}) & \sigma_{L2} (1 - p_{L2}) & \sigma_{L3} & 0 & 0 & 0 & 0 \\
  0 & 0 & 0 & \sigma_{J1} p_{J1} & 0 & 0 & 0 \\
  0 & 0 & 0 & \sigma_{J1} (1 - p_{J1}) \omega & \sigma_{J2} \omega & \sigma_{A_R} & 0 \\
  0 & 0 & 0 & 0 & 0 & 0 & \sigma_{A_T} 
\end{bmatrix} \begin{bmatrix}
L_1 \\
L_2 \\
L_3 \\
J_1 \\
J_2 \\
A_R \\
A_T \\
\end{bmatrix}(t)
$$

The parameters in this model are yearly survival probability $\sigma_{*}$, where the "\*" indicates a particular state variable, the probability that a female frog reproduces in a given year $p_F$, the number of eggs produced by a female frog in a year that successfully hatch $F$, the probability of a first year tadpole remaining as a tadpole $p_{L1}$, the probability of a second year tadpole remaining as a tadpole $p_{L2}$, and the probability of a first year juvenile remaining as a juvenile $p_{J1}$. The parameter $\omega$ gives the probability of a juvenile successfully recruiting to an adult.

In this model we ignore density-dependent recruitment because we were interested in the growth of the population from an initial reintroduction and whether this growth was sufficient to prevent extinction over 50 years following the introduction. We also did not directly consider the dynamics of Bd in this model. We made this decision because i) all translocated populations are infected with Bd at high prevalence ii) host density does not seem to play a significant role in multi-year Bd infection dynamics in this system. Thus, ignoring Bd infection dynamics and instead assuming all host vital rates are in the presence of high Bd prevalence significantly simplifies the model without much loss of realism.

**Model analysis**

**Incorporating yearly variability in vital rates**

We extracted yearly survival probabilities for translocated adults $\sigma_{A_T}$ and naturally recruited adults $\sigma_{A_R}$ from the mark-recapture model described elsewhere. While we observed yearly variability in adult survival *within* a population, the magnitude of this variability was small compared to *among* population variability. Thus, we did not include yearly within-population variability in adult survival in this analysis.

However, within a population there was substantial yearly variability in the successful recruitment of adults, greater than what we would expect from Poisson variability about a mean value. Therefore, we allowed for yearly variability in the probability of successful recruitment $\omega$ (more details below).

**Estimating model parameters**

The baseline parameter values for the model and how they were estimated are given in Table 1. Parameters $\sigma_{A_T}$ and $\sigma_{A_R}$ were extracted directly from our mark-recapture model described elsewhere. For populations where we had a sufficient number of pit-tagged, naturally recruited adults, we observed that $\sigma_{A_T}$ and $\sigma_{A_R}$ could be notably different, with $\sigma_{A_R} > \sigma_{A_T}$. For populations where we had not pit-tagged any naturally recruited adults, we could not directly estimate $\sigma_{A_R}$ and set $\sigma_{A_R} = \sigma_{A_T}$.

**Model analysis and simulation**

We performed four analyses on our model. First, we considered a deterministic version of our model with no yearly heterogeneity in probability of successful recruitment $\omega$ and calculated the predicted long-run growth rate $\lambda$ of a population for different values of $\sigma_{A_R}$ and $\omega$. We then fixed $\omega = 0.1$ and calculated the predicted growth rate of our six populations. Second, we performed a local sensitivity analysis on $\lambda$ with respect to parameters $\omega$, $\sigma_{A_R}$, and $F$ to determine how small changes in these parameters could influence the long-run deterministic growth rate of populations.

Third, we defined version of the model with demographic and environmental stochasticity, where environmental stochasticity was represented by among year variability in $\omega$. We used this model to simulate a one time introduction of 40 translocated adult frogs. We ran this simulation 1000 times for each population and computed the probability of a population becoming extinct after 50 years given the observed parameter values and environmental stochasticity in $\omega$. We varied the mean recruitment probability $\omega$ from 0.05 and 0.3 and drew values of $\omega$ each year from a beta distribution with a dispersion parameter of $\phi = 2$ (when $\omega = 0.5$ and $\phi = 2$ the beta distribution is uniform between 0 and 1).

Finally, we assessed whether our stochastic model could reproduce observed trajectories of population recovery. We focused on population 70550 because this was our longest mark-recapture time series for a translocated population and because this population shows evidence of recovery. We simulated our model for 16 years, repeating the simulation 50,000 times. For each run and each year, we drew $\omega$ from a uniform distribution between 0 and 1 (or a beta distribution with mean 0.5 and $\phi = 2$). Using Approximate Bayesian Computing and rejection sampling (Kosmala et al. 2015, Ecology), we identified the top 2% of trajectories (i.e., 1000 trajectories) that minimized the sum of squared errors between the observed and predicted data. The $\omega$ values associated with these "best" trajectories represented an approximate posterior distribution (Beaumont 2010). Using these best fit trajectories, we assessed whether our model could qualitatively describe the patterns of recovery in the observed data for population 70550.

@@

Science (impact factor = 63.714)

@@

Nature (impact factor = 69.8)

@@

PNAS (impact factor = 12.77)

@@

Nature Communications (impact factor = 17.694)

@@

    -   Papers published by the journal represent important advances of significance to specialists within each field. Open access.

@@

Science Advances (impact factor = 14.957)

-   Publishes impactful research papers and reviews in any area of science. Journal extends the capacity of *Science* magazine to identify and promote significant advances in science and engineering. Open access.

-   Research article

    -   Should present a major advance.

    -   Articles can be short of long format, up to 15,000 words, with up to 10 figures and/or tables.
